### 2025-05-03 lecturenotes_aima203/ch1_aima203.tex
general error

@@

\item Find the absolute error in the product $uv$, where $u=4.536$ and $v=1.32$, the numbers being correct up to the digits given. Find the relative error of the quotient $u/v$ as well.
\end{enumerate}

>>>>

\section{General Error Formula}
Let us consider a function $u$ that depends upon the variables: $x,y,z$, i.e $u=f(x,y,z)$. Let $\Delta x, \Delta y, \Delta z$ be the errors in $x$, $y$, and $z$, respectively. Then the total error in $u$ is $\Delta u$, which is approximated by $du$ as follows: We have the total derivative:
\begin{equation}
  \label{totalderive}
  du = \frac{\partial u}{\partial x} \,dx + \frac{\partial u}{\partial y} \,dy + \frac{\partial u}{\partial z} \,dz
\end{equation}
Approximating $dx$ by $\Delta x $, $dy$ by $\Delta y$ and $dz$ by $\Delta z$ we get,
\begin{equation}
  \label{generalerror}
  \Delta u \approx  \frac{\partial u}{\partial x} \,\Delta x + \frac{\partial u}{\partial y} \,\Delta y + \frac{\partial u}{\partial z} \,\Delta z
\end{equation}
The relation ~\ref{generalerror} gives the absolute general error formula. Then the relative error formula is given by $E_R = \frac{\Delta u}{u}$.

@@

\item Find the absolute error in the product $uv$, where $u=4.536$ and $v=1.32$, the numbers being correct up to the digits given. Find the relative error of the quotient $u/v$ as well.

\item Find the percentage error in $z=(1/8)\, xy^3$ when $x=3014 \pm 0.0016$ and $y=4.5 \pm 0.05$.
\item Prove that in a product of three nonzero numbers, the relative error does not exceed the sum of the relative errors of the numbers.
\end{enumerate}
